---
title: "Quiz3 Notebook"
output: html_notebook
---

Continuous Mixture:
```{r}
## simulate mixture of normal
set.seed(123)
S <- sample(c(1:3),10000,prob=c(0.2,0.5,0.3),replace=TRUE)
table(S)

x <- rep(NA,10000)
x[S==1] <- rnorm(length(which(S==1)),0,1)
x[S==2] <- rnorm(length(which(S==2)),-1,1)
x[S==3] <- rnorm(length(which(S==3)),2,1)
```
```{r}
hist(x)
```

Two ways of simulating possion with gamma prior of rate:
1.negative binomial
```{r}
library(MASS)
n <- 1000
p <- 0.4
r <- 10
mu <- p*r/(1-p)
sigma2 <- p*r/(1-p)^2
theta <- mu^2/(sigma2-mu) ## also equals to r, i.e, theta=r
x1 <- rnegbin(n,mu,theta)
```
2. possion+gamma
```{r}
## use the poisson-gamma formula
lambda1 <- rgamma(n,r,scale=p/(1-p))
x2 <- rpois(n,lambda1) #***
#*** is also equivalent to the following
x2 <- rep(NA,n)
for(i in 1:n){
  x2[i] <- rpois(1,lambda1[i])
}

qqplot(x1,x2)
abline(0,1)
```
eigendecomposition and cholesky
```{r}
my.Sigma <- matrix(c(2,1,1,1,2,1,1,1,2),nrow=3)
eigen(my.Sigma, symmetric = T)
L <- chol(my.Sigma)
L
t(L)%*%L
my.Sigma ==t(L)%*%L
```

Multivariate distribution
```{r}
genmultv.eigen <- function(n,mu,Sigma){
  d <- length(mu)
  ev <- eigen(Sigma,symmetric=TRUE)
  eigen.values <- ev$values
  P <- ev$vectors
  Delta <- diag(sqrt(eigen.values))
  Q <- P%*%Delta%*%t(P)
  Z <- matrix(rnorm(n*d),nrow=n)
  X <- Z%*%Q+matrix(rep(1,n),ncol=1)%*%t(mu)
  return(X)
}

my.mu <- c(0,0,0)
my.Sigma <- matrix(c(2,1,1,1,2,1,1,1,2),nrow=3)
my.X <- genmultv.eigen(10000,my.mu,my.Sigma)

colMeans(my.X)
apply(my.X,2,mean)
cov(my.X)
```
Cholesky 
```{r}
### generate multivariate normal using Cholesky decomposition ###

genmultv.chol <- function(n,mu,Sigma){
  d <- length(mu)
  Q <- chol(Sigma)
  Z <- matrix(rnorm(n*d),nrow=n)
  X <- Z%*%Q+matrix(mu, byrow = T, nrow = n, ncol = length(mu))
  return(X)
}

my.mu <- c(0,0,0)
my.Sigma <- matrix(c(2,1,1,1,2,1,1,1,2),nrow=3)
my.Xchol <- genmultv.chol(10000,my.mu,my.Sigma)

colMeans(my.Xchol)
##
apply(my.Xchol,2,mean)
cov(my.Xchol)
```
matrix(my.mu, nrow = 10, byrow = T)
```{r}
matrix(my.mu, nrow = 10, ncol =3, byrow = F)
```
Generalized inverse

```{r}
x <- matrix(c(2,4,1,2), nrow = 2)
y <- svd(x)
y
```
Regualr inverse when the matrix is square and full rank
```{r}
x <- matrix(c(1,2,3,4), nrow = 2)
y <- solve(x)
x%*%y
```
solve() function has many other usage, should be familarize it later on.

Smooting splines
```{r}
#loading the Splines Packages
require(splines)
#ISLR contains the Dataset
require(ISLR)
attach(Wage) #attaching Wage dataset
?Wage #for more details on the dataset
agelims<-range(age)
#Generating Test Data
age.grid<-seq(from=agelims[1], to = agelims[2])
```
fit a Cubic Spline with 3 Knots (cutpoints) 

```{r}
#3 cutpoints at ages 25 ,50 ,60
fit<-lm(wage ~ bs(age,knots = c(25,40,60)),data = Wage )
summary(fit)
```

